---
title: "vic_analysis"
author: "Rachel"
date: "11/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r}
#install.packages("mapplots")
#install.packages("raster")
library(mapplots)
library(raster)
library(sf)
library(lubridate)
library(tidyverse)
```


###load data from "./data_do_not_push"

```{r}
getwd()
load("../data_do_not_push/VIC_monthly_Precip_PET_Temp.Rdata")
```

#subtract baseline from hot/dry to determine change for every timestep in every gridcell
```{r}

diff_precip <- vic_monthly$hotdry_precip - vic_monthly$base_precip
diff_temp <- vic_monthly$hotdry_temp - vic_monthly$base_temp
diff_pet <- vic_monthly$hotdry_pet - vic_monthly$base_pet


```

#add back in date column 
```{r}
Datecol <- vic_monthly$base_precip$date
diff_precip$date <- Datecol
diff_temp$date <- Datecol
diff_pet$date <- Datecol
```

#average through time (to produce one value for each grid cell)
```{r}
#annual_avg_precip <- colMeans(diff_precip[,-1])
annual_avg_temp <- colMeans(diff_temp[,-1])
annual_avg_pet <- colMeans(diff_pet[,-1])
```


#create raster

rasterFromXYZ(xyz, res=c(NA,NA), crs="", digits=5)
```{r}
x <- vic_monthly$lon
y <- vic_monthly$lat

#precip
#xyz <- data.frame(x,y,annual_avg_precip)
#
#rst <- rasterFromXYZ(xyz)
#crs(rst) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #confirm crs
#crs(rst)

#writeRaster(rst, './output/precip_chg.tif')

#temp
xyz <- data.frame(x,y,annual_avg_temp)

rst <- rasterFromXYZ(xyz)
crs(rst) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #confirm crs
crs(rst)

writeRaster(rst, '../output/temp_chg.tif')

#pet
xyz <- data.frame(x,y,annual_avg_pet)

rst <- rasterFromXYZ(xyz)
crs(rst) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #confirm crs
crs(rst)

writeRaster(rst, '../output/pet_chg.tif')
```


#change factor for precip
```{r}
chg_factor_precip <- vic_monthly$hotdry_precip[-1]/vic_monthly$base_precip[-1]


#month change factor for precip to compare to vol2 section 14 monthly change factors
precip_monthly <- chg_factor_precip %>%
  mutate(date = ymd(Datecol)) %>%
  gather(key = "gridcell", value = "value", -date) %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarise(avg_precip = mean(value, na.rm = TRUE)) 

#average through time (to produce one value for each grid cell)
chg_factor_precip[is.na(chg_factor_precip)] <- 0
chg_factor_avg_precip <- colMeans(chg_factor_precip)


#write raster for change factor 
xyz <- data.frame(x,y,chg_factor_avg_precip)

rst <- rasterFromXYZ(xyz)
crs(rst) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #confirm crs
crs(rst)
writeRaster(rst, '../output/precip_chg_factor.tif')
```














#create grid
###did not work--"could not find function geoTiff"
make.grid(x, y, z, byx , byy , xlim, ylim, fun = function(x) sum(x, na.rm = T))

Arguments
x  a vector of x-coordinates (longitude)
y  a vector of y-coordinates (latitude; same length as x)
z  a vector of values; same length as x
byx, byy  the size of the grid cells on the x and y scale
xlim, ylim  the x and y limits of the grid. Note that the origin of the grid depends on xlim[1] and ylim[1], these values will be taken as the mid-point of the bottom-left grid-cell. 
fun  a function to be applied to z. Defaults to sum.
```{r}
minlon <- min(vic_monthly$lon)
maxlon <- max(vic_monthly$lon)
minlat <- min(vic_monthly$lat)
maxlat <- max(vic_monthly$lat)
xlim <- c(minlon, maxlon)
ylim <- c(minlat, maxlat)
x <- vic_monthly$lon
y <- vic_monthly$lat
byx <- 1
byy <- 0.5

grd <- make.grid(x, y, annual_avg_precip, byx, byy, xlim, ylim)
draw.grid(grd,breaks)

geoTiff(grd, "./output/test.tiff")
mapplots::geoTiff(grd, "./output/test.tiff")
rgdal::geoTiff(grd, "./output/test.tiff")

```

